---
title: "hw4_hw2570"
author: "Hongbo Wang hw2570 Sec D01"
date: "October 24, 2016"
output: md_document
---
### i.
```{r}
rm(list = ls())
gmp <- read.table("~/Desktop/R data/gmp.txt", head = TRUE)
gmp$pop <- round(gmp$gmp / gmp$pcgmp)
names(gmp)

plot(y = gmp$pcgmp, x = gmp$pop, log = "x")
curve(6611*x^(1/8), add = TRUE, col = "blue")     # b0 = 6611, b1 = 1/8
curve(6611*x^(0.1), add = TRUE, col = "red")
curve(6611*x^(0.15), add = TRUE, col = 3)
legend("bottomright", legend = c("b1 = 1/8", "b1 = 0.1", "b1 = 0.15"), 
       col=c("blue", "red", 3), pch = 1)
```

### ii.
```{r}
mse <- function(r = c(b0, b1), X = gmp$pop, Y = gmp$pcgmp){
  sse <- sum( ( Y - r[1]*(X^r[2]) ) ^2 )
  mse <- sse / length(X)
  return(mse) 
}
mse(c(6611, 0.15))
mse(c(5000, 0.1))
```

### iii.
```{r, warning = FALSE}
nlm(mse, c(beta0 = 6611, beta1 = 1/8))
nlm(mse, c(beta0 = 6611, beta1 = 0.1))
nlm(mse, c(beta0 = 6611, beta1 = 0.15))
nlm(mse, c(beta0 = 6611, beta1 = 1/8))$estimate
```

### iv.

```{r, warning = FALSE}
plm <- function(b0, b1, X = gmp$pop, Y = gmp$pcgmp){
  est_b0 <- nlm(mse, c(b0, b1))$estimate[1]
  est_b1 <- nlm(mse, c(b0, b1))$estimate[2]
  est_MSE <- nlm(mse, c(b0, b1))$minimum
  return(c(est_b0, est_b1, est_MSE))
}
plm(6611, 0.15)
plm(5000, 0.10)
```

We can see that the data parameter estimated when starting from *β0 = 6611*, *β1 = 0.15* and from *β0 = 5000* and *β1 = 0.10* are not the same. The reason is most data are located in the range (5e+04, 2e+07), which is really far from the origin. So the affecting from *β0*, the *y-interval* is negligible. Different *β0* will return to a correspondent *β1*. 

In this question, *β0 = 6611*, *β1 = 0.15* returns $6.185706e+07$ and *β0 = 5000* and *β1 = 0.10* returns $6.252148e+07$. So *β0 = 6611*, *β1 = 0.15* have a smaller *MSE*.

### v.
### (a).
```{r}
xbar <- mean(gmp$pcgmp); std_x <- sd(gmp$pcgmp)
n <- length(gmp$pcgmp)
(sem <- std_x / sqrt(n))
```

### (b).
```{r}
indices <- function(ind = vector(mode = "numeric"), 
                    X = gmp$pcgmp){
  mean_ind <- mean(X[ind])
  return(mean_ind)  
}
```

### (c).
```{r}
bootstrap_means <- vector(length = 100)
for(i in 1:100){
  index <- sample(1:length(gmp$pcgmp), 100, replace = TRUE)
  bootstrap_means[i] <- indices(index)
}
```

### (d).
```{r}
(sem_bootstrap <- sd(bootstrap_means))
```

The value is not closed to the value in part (a). The reason for that is because the size of B is too small, so the distribution of bootstrap mean does not fit the normal distribution very well. We can see that from the picture. 

```{r}
hist(bootstrap_means)
```


### (vi).
```{r, warning = FALSE}
plm <- function(b0, b1, X, Y){
  mse <- function(r = c(b0, b1)){
    sse <- sum( ( Y - r[1]*(X^r[2]) ) ^2 )
    mse <- sse / length(X)
    return(mse)}
  est_b0 <- nlm(mse, c(b0, b1))$estimate[1]
  est_b1 <- nlm(mse, c(b0, b1))$estimate[2]
  est_MSE <- nlm(mse, c(b0, b1))$minimum
  return(c(est_b0, est_b1, est_MSE))
}

plm_bootstrap <- function(b0, b1, X = gmp$pop, Y = gmp$pcgmp, B = 100){
  b0_boot <- vector(length = B)
  b1_boot <- vector(length = B)
  for(i in 1:B){
    index <- sample(1:length(X), B , replace = TRUE)  
    b0_boot[i] <- plm(b0, b1, X[index], Y[index])[1]
    b1_boot[i] <- plm(b0, b1, X[index], Y[index])[2]
  }
  return(c(sd(b0_boot), sd(b1_boot)))
}

plm_bootstrap(6611, 0.15)
plm_bootstrap(5000, 0.10)
```

### (vii).

```{r, warning = FALSE}
gmp_2013 <- read.table("~/Desktop/R data/gmp-2013.txt", head = TRUE)
gmp_2013$pop <- round(gmp_2013$gmp / gmp_2013$pcgmp)

plm_bootstrap <- function(b0, b1, X = gmp_2013$pop, Y = gmp_2013$pcgmp, B = 100){
  b0_boot <- vector(length = B)
  b1_boot <- vector(length = B)
  for(i in 1:B){
    index <- sample(1:length(X), B , replace = TRUE)  
    b0_boot[i] <- plm(b0, b1, X[index], Y[index])[1]
    b1_boot[i] <- plm(b0, b1, X[index], Y[index])[2]
  }
  return(c(sd(b0_boot), sd(b1_boot)))
}

plm(6611, 0.15, gmp_2013$pop, gmp_2013$pcgmp)
plm_bootstrap(6611, 0.15)
```
As we see in the result, the standard deviation of the estimated parameters in the data set *gmp* and *gmp_2013* are closed.
The estimated parameter we got from previous question are $6611.0000000$ and $0.1263177$, and this time we have the estimated parameter $6611$ and $0.1433688$. Since they are really closed, the parameters of the model does not changed significantly.





