---
title: "hw7_hw2570"
author: "Hongbo Wang hw2570"
date: "November 14, 2016"
output: md_document
---
```{r}
rm(list = ls())
Data <- read.csv("~/Desktop/R data/moretti.csv", header = TRUE)
```

### Part 1
### i.
```{r}
poisLoglik <- function(lambda, data){
  value <- sum( log(((lambda^(data)) * exp(-lambda))/factorial(data)) )
  return(value) 
}

data = c(1, 0, 0, 1, 1)
poisLoglik(1, data)
```

### ii.
```{r}
count_new_genres <- function(year){
  if(length(which(Data$Begin != year)) == 44){return(0)}
  else{return(length(which(Data$Begin == year)))}
}

count_new_genres(1803)
count_new_genres(1850)
```

### iii.
```{r}
new_genres <- rep(NA)
for(i in 1740:1900){
  new_genres[i - 1739] <- count_new_genres(i)
}
(pos_1803 <- 1803 - 1739) ; (pos_1850 <- 1850 - 1739)
new_genres[pos_1803]; new_genres[pos_1850]
```

### iv.
```{r}
#poisLoglik_plot <- function(lambda, data = new_genres){
#  return(sapply(seq(from = 0, to = 5, by = 0.001), poisLoglik, data = new_genres))}
#curve(poisLoglik_plot(lambda = x), from = 0, to = 5)
x <- seq(from = 0, to = 5, by = 0.001)
y <- sapply(seq(from = 0, to = 5, by = 0.001), poisLoglik, data = new_genres)
plot(seq(from = 0, to = 5, by = 0.001), y, type = 'l')
index <- which(y == max(y))
lines(c(x[index], x[index]), c(-800, max(y) + 50), col = 2, lty = 2)
x[index]
```

### v.
```{r, warning = FALSE}
poisLoglik <- function(lambda, data = new_genres){
  value <- sum( log(((lambda^(data)) * exp(-lambda))/factorial(data)) )
  return(-value) 
}
nlm(poisLoglik, p = 1)$estimate
```

The MLE estimation is same as the plot we have at last question.

### vi.
```{r}
intergenre_intervals <- diff(Data$Begin)
head(intergenre_intervals)
mean_ii <- mean(intergenre_intervals)
std <- sd(intergenre_intervals)
coefficient_of_variation <- std / mean_ii
data.frame(mean_ii, std, coefficient_of_variation)
```

### vii.
### a.
```{r}
sim_interval <- function(new_genres){
  new_year <- which(new_genres != 0)
  i_i <- c(NA)
  for(i in 2 : length(new_year)){
    i_i[i-1] <- new_year[i] - new_year[i-1]
  }
  return(i_i)
}
head(sim_interval(new_genres)); head(intergenre_intervals)
```

### b.
```{r}
sim_poisson <- function(num_year, lambda){
  sim_pois <- rpois(num_year, lambda)
  i_i <- sim_interval(sim_pois)
  cov <- sd(i_i) / mean(i_i)
  return(c(cov, i_i))
}
sim_1 <- sim_poisson(161, 0.273)
sim_1[1]
mean(sim_1[-1])
```

The mean of the intervals does between 3 and 4.

### viii.
```{r}
sim_cov <- c(NA)
for(i in 1:100000){
  sim_cov[i] <- sim_poisson(161, 0.273)[1]
}
length(sim_cov[sim_cov > coefficient_of_variation]) / length(sim_cov)
```

### ix.
We can see only a small portion of simulations runs have a higher coefficient of variation than Moretti’s data, which means genres did appear randomly, at a constant rate – a Poisson process. So the conjecture that that genres tend to appear together in bursts is true.




